\textbf{Supplementary Information}

### 2020-02-14 submission/supplementary.tex
Added introduction and supp informations.

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{\textbf{Dynamics of Epiphytic Microbial Communities on Marine Macrophyte
Surfaces}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

\usepackage[labelsep=period, labelfont=bf]{caption}
\renewcommand{\thefigure}{\arabic{figure}}
\captionsetup{justification=raggedright,singlelinecheck=false}

@@

\usepackage{xr}
\externaldocument[supp-]{supplementary}

@@

2. University of Vienna, Department of Limnology and Bio-Oceanography,
Althanstraße 14, Vienna, Austria \newpage \linenumbers
\sisetup{mode=text} \setlength\parindent{24pt}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Marine macrophytes (seagrasses and macroalgae) are important ecosystem
engineers that form close associations with microorganism belonging to
all three domains of life (Egan \emph{et al.}, 2013; Tarquinio \emph{et
al.}, 2019). Microbes can live within macrophyte tissue as endophytes or
can form epiphytic communities on surfaces of leaves and thalli (Egan
\emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
al.}, 2019). Epiphytic and endophytic microbial communities form a close
functional relationship with the macrophyte host. It was proposed that
this close relationship constitutes a holobiont, an integrated community
where the macrophyte organism and its symbiotic partners support each
other (Margulis, 1991; Egan \emph{et al.}, 2013; Tarquinio \emph{et
al.}, 2019).

Biofilms formed from microbial epiphytes can contain diverse taxonomic
groups and harbor cell densities from 10\textsuperscript{2} to
10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong \emph{et
al.}, 2000; Bengtsson \emph{et al.}, 2010; Burke \emph{et al.}, 2011b).
In such an environment a number of positive and negative interactions
between the macrophyte and colonizing microorganisms have been described
(Egan \emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio
\emph{et al.}, 2019). Macrophytes can promote growth of associated
microbes by nutrient exudation, while in return microorganisms may
support macrophyte performance through improved nutrient availability,
phytohormone production and protection form toxic compounds, oxidative
stress, biofouling organisms and pathogens (Egan \emph{et al.}, 2013;
Hollants \emph{et al.}, 2013; Tarquinio \emph{et al.}, 2019). Beside
this positive interactions, macrophytes can negatively impact the
associated microbes such as pathogenic bacteria by producing reactive
oxygen species and secondary metabolites (Egan \emph{et al.}, 2013;
Hollants \emph{et al.}, 2013; Tarquinio \emph{et al.}, 2019).

All these ecological roles are carried out by a taxonomically diverse
community of microorganisms. At higher taxonomic ranks a core set of
macrophyte epiphytic taxa was described consisting mainly of
\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria},
\emph{Desulfobacterota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
\emph{Actinobacteriota}, \emph{Firmicutes}, \emph{Planctomycetota},
\emph{Chloroflexi} and \emph{Verrucomicrobiota} (Crump and Koch, 2008;
Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). In contrast,
at lower taxonomic ranks host specific microbial communities were
described (Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.},
2016). Recently, it was shown that even different morphological niches
within the same alga had a higher influence on bacterial community
variation than biogeography or environmental factors (Morrissey \emph{et
al.}, 2019). While there is high community variation between host
species is was observed that the majority of metagenomes determined
functions were conserved both between host species and individuals
(Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). This
discrepancy between taxonomic and functional composition could be
explained by the lottery hypothesis. It postulates that an initial
random colonization step is performed from a set of functionally
equivalent taxonomic groups resulting in taxonomically different
epiphytic communities sharing a core set of functional genes (Burke
\emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). In addition,
some of the variation in the observed data could be attributed to
different techniques used in various studies, such as different
protocols for epiphytic cell detachment and/or DNA isolation, as no
standard protocol to study epiphytic communities was established
(Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.}, submitted).

The majority of studies describing macrophyte epiphytic communities did
not encompass seasonal changes (Crump and Koch, 2008; Lachnit \emph{et
al.}, 2009; Burke \emph{et al.}, 2011b; Roth-Schulze \emph{et al.},
2016; Ugarelli \emph{et al.}, 2019). In addition, if seasonal changes
were taken into account low temporal frequency and/or methodologies that
do not allow for high taxonomic resolution were used (Tujula \emph{et
al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013).
In the present study we describe the seasonal dynamics of bacterial and
archaeal communities on the surfaces of the seagrass \emph{Cymodocea
nodosa} and siphonous macroalgae \emph{Caulerpa cylindracea} determined
on a mostly monthly scale. Bacterial and archaeal epiphytes were sampled
in a meadow of \emph{Cymodocea nodosa} invaded by the invasive
\emph{Caulerpa cylindracea} and in a locality of only \emph{Caulerpa
cylindracea} located in the proximity of the meadow. In addition, for
comparison the community of the surrounding seawater was characterized.

\newpage

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\hypertarget{sampling}{%
\subsubsection{Sampling}\label{sampling}}

Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
nodosa} meadow located in the proximity of the village of Funtana
(\ang{45;10;39} N, \ang{13;35;42} E). Thalli of \emph{Caulerpa
cylindracea} were sampled in the same \emph{Cymodocea nodosa} invaded
meadow in Funtana and on a locality of only \emph{Caulerpa cylindracea}
located close to the invaded meadow. Sampling of leaves and thalli was
performed approximately monthly from December 2017 to October 2018.
Leaves and thalli were collected by diving and transported to the
laboratory in containers placed on ice and filled with site seawater.
Upon arrival to the laboratory, \emph{Cymodocea nodosa} leaves were cut
into sections of 1 -- 2 \si{\cm}, while \emph{Caulerpa cylindracea}
thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli
were washed three times with sterile artificial seawater (ASW) to remove
loosely attached microbial cells. Surrounding seawater was collected in
10 \si{\l} containers by diving and transported to the laboratory where
the whole container volume was filtered through a 20 \si{\um} net. The
filtrate was further sequentially filtered through 3 \si{\um} and 0.2
\si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using
a peristaltic pump. Filters were briefly dried at room temperature and
stored at \num{-80} \si{\degreeCelsius}. Seawater samples were also
collected approximately monthly from July 2017 to October 2018.

\hypertarget{dna-isolation}{%
\subsubsection{DNA Isolation}\label{dna-isolation}}

DNA from surfaces of \emph{Cymodocea nodosa} and \emph{Caulerpa
cylindracea} was isolated using a previously modified and adapted
protocol that allows a selective epiphytic DNA isolation (Massana
\emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
and thalli are incubated in a lysis buffer and treated with lysozyme and
proteinase K. Following the incubations, the mixture containing lyzed
epiphytic cells is separated from leaves and thalli and extracted using
a phenol-chloroform procedure. Finally, the extracted DNA is
precipitated using isopropanol. DNA from seawater picoplankton was
isolated from 0.2 \si{\um} polycarbonate filters according to (Massana
\emph{et al.}, 1997) with a slight modification. Following the
phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium
acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of
chilled isopropanol, incubating the mixtures overnight at \num{-20}
\si{\degreeCelsius} and centrifuging at 20,000 × g and 4
\si{\degreeCelsius} for 21 \si{\minute}. The pellet was washed twice
with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5
\si{\minute}. Dried pellets were resuspended in 50 -- 100 \si{\ul} of
deionized water.

\hypertarget{illumina-16s-rrna-sequencing}{%
\subsubsection{Illumina 16S rRNA
Sequencing}\label{illumina-16s-rrna-sequencing}}

Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
described previously (Korlević \emph{et al.}, submitted). The V4 region
of the 16S rRNA gene was amplified using a two-step PCR procedure. In
the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Purified PCR products were
sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried,
Germany. Before sequencing, the second PCR amplification of the two-step
PCR procedure was performed at IMGM using primers targeting the tagged
region incorporated in the first PCR. In addition, these primers
contained adapter and sample-specific index sequences. Beside samples, a
positive and negative control for each sequencing batch were sequenced.
A negative control was comprised of PCR reactions without DNA template,
while for a positive controls a mock community composed of evenly mixed
DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
USA) was used. The sequences obtained in this study have been submitted
to the European Nucleotide Archive (ENA) under accession numbers
\textbf{TO BE ADDED LATER!}.

\hypertarget{sequence-analysis}{%
\subsubsection{Sequence Analysis}\label{sequence-analysis}}

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
was done using R (version 3.6.0) (R Core Team, 2019), package vegan
(version 2.5-6) (Oksanen \emph{et al.}, 2019), package tidyverse
(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
packages (Xie, 2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie
\emph{et al.}, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
detailed analysis procedure including the R Markdown file for this paper
are available as a GitHub repository (\textbf{TO BE ADDED LATER!}).
Based on the ATCC MSA-1002 mock community included in the analysis an
average sequencing error rate of 0.01 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
2016). In addition, the negative controls processed together with the
samples yielded on average only 2 sequences after sequence quality
curation.

\hypertarget{results}{%
\subsection{Results}\label{results}}

Chloroplast sequences were following the seasonal pattern
(\autoref{supp-chloroplast}, \autoref{pcoa}).

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019) rmarkdown: Dynamic Documents for R.

\leavevmode\hypertarget{ref-Apprill2015}{}%
Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
revision to V4 region SSU rRNA 806R gene primer greatly increases
detection of SAR11 bacterioplankton. \emph{Aquatic Microbial Ecology}
\textbf{75}: 129--137.

\leavevmode\hypertarget{ref-Armstrong2000}{}%
Armstrong, E., Rogerson, A., and Leftley, J.W. (2000) The Abundance of
Heterotrophic Protists Associated with Intertidal Seaweeds.
\emph{Estuarine, Coastal and Shelf Science} \textbf{50}: 415--424.

\leavevmode\hypertarget{ref-Bengtsson2010}{}%
Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
bacterial biofilms on the kelp Laminaria hyperborea. \emph{Aquatic
Microbial Ecology} \textbf{60}: 71--83.

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011a) Bacterial community assembly based on functional genes rather
than species. \emph{Proceedings of the National Academy of Sciences of
the United States of America} \textbf{108}: 14288--14293.

\leavevmode\hypertarget{ref-Burke2011b}{}%
Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
(2011b) Composition, uniqueness and variability of the epiphytic
bacterial community of the green alga Ulva australis. \emph{The ISME
journal} \textbf{5}: 590--600.

\leavevmode\hypertarget{ref-Caporaso2012}{}%
Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
analysis on the Illumina HiSeq and MiSeq platforms. \emph{The ISME
Journal} \textbf{6}: 1621--1624.

\leavevmode\hypertarget{ref-Crump2008}{}%
Crump, B.C. and Koch, E.W. (2008) Attached bacterial populations shared
by four species of aquatic angiosperms. \emph{Applied and environmental
microbiology} \textbf{74}: 5948--5957.

\leavevmode\hypertarget{ref-Egan2013}{}%
Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
interactions. \emph{FEMS microbiology reviews} \textbf{37}: 462--476.

\leavevmode\hypertarget{ref-Hollants2013}{}%
Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013) What
we can learn from sushi: A review on seaweed-bacterial associations.
\textbf{83}: 1--16.

\leavevmode\hypertarget{ref-Korlevic_in_press}{}%
Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
Selective DNA and Protein Isolation from Marine Macrophyte Surfaces.

\leavevmode\hypertarget{ref-Kozich2013}{}%
Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
curation pipeline for analyzing amplicon sequence data on the MiSeq
Illumina sequencing platform. \emph{Applied and environmental
microbiology} \textbf{79}: 5112--5120.

\leavevmode\hypertarget{ref-Lachnit2009}{}%
Lachnit, T., Blümel, M., Imhoff, J.F., and Wahl, M. (2009) Specific
epibacterial communities on macroalgae: Phylogeny matters more than
habitat. \emph{Aquatic Biology} \textbf{5}: 181--186.

\leavevmode\hypertarget{ref-Lachnit2011}{}%
Lachnit, T., Meske, D., Wahl, M., Harder, T., and Schmitz, R. (2011)
Epibacterial community patterns on marine macroalgae are host-specific
but temporally variable. \emph{Environmental Microbiology} \textbf{13}:
655--665.

\leavevmode\hypertarget{ref-Margulis1991}{}%
Margulis, L. (1991) Symbiogenesis and symbionticism. In, Margulis,L. and
Fester,R. (eds), \emph{Symbiosis as a source of evolutionary
innovation}. Cambridge, Massachusetts: The MIT Press, pp. 1--14.

\leavevmode\hypertarget{ref-Massana1997}{}%
Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
Vertical distribution and phylogenetic characterization of marine
planktonic Archaea in the Santa Barbara Channel. \emph{Applied and
environmental microbiology} \textbf{63}: 50--56.

\leavevmode\hypertarget{ref-Miranda2013}{}%
Miranda, L.N., Hutchison, K., Grossman, A.R., and Brawley, S.H. (2013)
Diversity and Abundance of the Bacterial Community of the Red Macroalga
Porphyra umbilicalis: Did Bacterial Farmers Produce Macroalgae?
\emph{PLoS ONE} \textbf{8}: e58269.

\leavevmode\hypertarget{ref-Morrissey2019}{}%
Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
Disentangling the influence of environment, host specificity and thallus
differentiation on bacterial communities in siphonous green seaweeds.
\emph{Frontiers in Microbiology} \textbf{10}:

\leavevmode\hypertarget{ref-Neuwirth2014}{}%
Neuwirth, E. (2014) RColorBrewer: ColorBrewer Palettes.

\leavevmode\hypertarget{ref-Oksanen2019}{}%
Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
McGlinn, D., et al. (2019) vegan: Community Ecology Package.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
matters: assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environmental Microbiology} \textbf{18}: 1403--1414.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013) The SILVA ribosomal RNA gene database project: improved
data processing and web-based tools. \emph{Nucleic acids research}
\textbf{41}: D590--6.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019) R: A Language and Environment for Statistical
Computing, Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Roth-Schulze2016}{}%
Roth-Schulze, A.J., Zozaya-Valdés, E., Steinberg, P.D., and Thomas, T.
(2016) Partitioning of functional and taxonomic diversity in
surface-associated microbial communities. \emph{Environmental
microbiology} \textbf{18}: 4391--4402.

\leavevmode\hypertarget{ref-Schloss2016}{}%
Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
(2016) Status of the Archaeal and Bacterial Census: an Update.
\emph{mBio} \textbf{7}: e00201--16.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
Hollister, E.B., et al. (2009) Introducing mothur: open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Applied and environmental
microbiology} \textbf{75}: 7537--7541.

\leavevmode\hypertarget{ref-Tarquinio2019}{}%
Tarquinio, F., Hyndes, G.A., Laverock, B., Koenders, A., and Säwström,
C. (2019) The seagrass holobiont: Understanding seagrass-bacteria
interactions and their role in seagrass ecosystem functioning.
\emph{FEMS Microbiology Letters} \textbf{366}:

\leavevmode\hypertarget{ref-Tujula2010}{}%
Tujula, N.A., Crocetti, G.R., Burke, C., Thomas, T., Holmström, C., and
Kjelleberg, S. (2010) Variability and abundance of the epiphytic
bacterial community associated with a green marine Ulvacean alga.
\emph{The ISME Journal} \textbf{4}: 301--311.

\leavevmode\hypertarget{ref-Ugarelli2019}{}%
Ugarelli, K., Laas, P., and Stingl, U. (2019) The microbial communities
of leaves and roots associated with turtle grass (Thalassia testudinum)
and manatee grass (syringodium filliforme) are distinct from seawater
and sediment communities, but are similar between species and sampling
sites. \emph{Microorganisms} \textbf{7}:

\leavevmode\hypertarget{ref-Wickham2019}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
R., et al. (2019) Welcome to the tidyverse. \emph{Journal of Open Source
Software} \textbf{4}: 1686.

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
Raton, Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
\{R\}. In, Stodden,V., Leisch,F., and Peng,R.D. (eds),
\emph{Implementing reproducible computational research}. Chapman;
Hall/CRC.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
Generation in R.

\leavevmode\hypertarget{ref-Xie2019}{}%
Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
30--32.

\leavevmode\hypertarget{ref-Xie2020}{}%
Xie, Y. (2020) tinytex: Helper Functions to Install and Maintain 'TeX
Live', and Compile 'LaTeX' Documents.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
Definitive Guide, Boca Raton, Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-Yilmaz2014}{}%
Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
D643--8.

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe
Syntax.

\newpage 
\setlength\parindent{0pt}

\hypertarget{figure-captions}{%
\subsection{Figure Captions}\label{figure-captions}}

\textbf{\autoref{shared}.} \nameref{shared}

\textbf{\autoref{pcoa}.} \nameref{pcoa}

\textbf{\autoref{community}.} \nameref{community}

\textbf{\autoref{cyano}.} \nameref{cyano}

\textbf{\autoref{bactero}.} \nameref{bactero}

\textbf{\autoref{alpha}.} \nameref{alpha}

\textbf{\autoref{gamma}.} \nameref{gamma}

\textbf{\autoref{desulfo}.} \nameref{desulfo}

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

@@

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/seasonal_shared}

@@

\caption{Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{shared}}\label{fig:unnamed-chunk-2}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{../results/figures/pcoa_figure}

@@

\caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater. Samples from the same environment or same season are labeld in different colors. The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-3}
\end{figure}

@@

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/community_bar_plot}

@@

\caption{Taxonomic classification and relative contribution of the most abundant bacterial and archaeal sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{community}}\label{fig:unnamed-chunk-4}
\end{figure}

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/cyanobacteria_bar_plot}

\caption{Taxonomic classification and relative contribution of the most abundant cyanobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater. The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative\label{cyano}}\label{fig:unnamed-chunk-5}
\end{figure}

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/bacteroidota_bar_plot}

\caption{Taxonomic classification and relative contribution of the most abundant sequences within the \textit{Bacteroidota} on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater. The proprotion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative\label{bactero}}\label{fig:unnamed-chunk-6}
\end{figure}

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/alphaproteobacteria_bar_plot}

\caption{Taxonomic classification and relative contribution of the most abundant alphaproteobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater. The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative\label{alpha}}\label{fig:unnamed-chunk-7}
\end{figure}

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/gammaproteobacteria_bar_plot}

\caption{Taxonomic classification and relative contribution of the most abundant gammaproteobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater. The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative\label{gamma}}\label{fig:unnamed-chunk-8}
\end{figure}

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/desulfobacterota_bar_plot}

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Information},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{\textbf{Supplementary Information}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{\textbf{Dynamics of Epiphytic Microbial Communities on Marine Macrophyte
Surfaces}}
  \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

\usepackage[labelsep=period, labelfont=bf]{caption}
\renewcommand{\thefigure}{S\arabic{figure}}
\captionsetup{justification=raggedright,singlelinecheck=false}

@@

@@

2. University of Vienna, Department of Limnology and Bio-Oceanography,
Althanstraße 14, Vienna, Austria

\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}

\hypertarget{supplementary-figures}{%
\subsection{Supplementary Figures}\label{supplementary-figures}}

@@

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/rarefaction}

@@

\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
\end{figure}

@@

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/calculators}

@@

\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
\end{figure}

@@

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/chloroplast_bar_plot}

@@

\caption{Relative contribution of chloroplast sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
\end{figure}
